---
title: "7005_HW8"
author: "Youngmin Choi"
date: "2024-03-14"
output: html_document
---
```{r setup}
echo=TRUE
message=FALSE
warning= FALSE
library(tidyverse)     
library(RPostgres)     
library(connections)   
library(DBI)
library(yaml)
library(dplyr)
library(ggplot2)
```

## Objective
In this task, multiple connections were generated to access databases from different sources in order to analyze the data of patients with different conditions. 


## Step 1: Connect to databases.


### 1) Define function to connect to multiple databases. 

A function was created to connect to multiple databases based on the connection information contained in the config_file yaml file. This was done to remove any personal secure information from the code and to connect to multiple connections without having to include more codes into the notebook. 
```{r}
establish_connections <- function(config_file) {
  config <- yaml::read_yaml(config_file)
  connections <- list()
  
  for (db_name in names(config$databases)) {
    db_config <- config$databases[[db_name]]
    print(paste("Connecting to", db_name))
    con <- connection_open(RPostgres::Postgres(),
                           host = db_config$host,
                           dbname = db_config$dbname,
                           user = db_config$user,
                           password = db_config$password)
    print(paste("Connection to", db_name, "established"))
    connections[[db_name]] <- con
  }
  
  return(connections)
}

```



### 2) Connect to databases & Load tables. 

Multiple connections made simultaneously while iterating the function defined above. 
```{r}
connections <- establish_connections("config.yaml")
connections
```

Each individual connections assigned to each variable to extract information individually. 
```{r}
db1_con<-connections$db1
db2_con<-connections$db2
db3_con<-connections$db3
db4_con<-connections$db4
db5_con<-connections$db5
db6_con<-connections$db6
db7_con<-connections$db7
db8_con<-connections$db8
db9_con<-connections$db9
```

Each tables were loaded separately in order to assess the table structure since each table may have different features. 

```{r}
pneumonia<-dbGetQuery(db1_con, "SELECT * FROM pneumonia")
pneumonia <- pneumonia %>% filter (is.na(deathdate))
head(pneumonia)
```

```{r}
viral_sinusitis<-dbGetQuery(db2_con, "SELECT * FROM viral_sinusitis")
viral_sinusitis <- viral_sinusitis %>% filter (is.na(deathdate))
head(viral_sinusitis)
```

```{r}
chronic_sinusitis<-dbGetQuery(db3_con, "SELECT * FROM Chronic_sinusitis")
chronic_sinusitis <- chronic_sinusitis %>% filter (is.na(deathdate)) %>% filter (birthdate <=Sys.Date())
head(chronic_sinusitis)
```


```{r}
osteoporosis<-dbGetQuery(db4_con, "SELECT * FROM osteoporosis")
osteoporosis <- osteoporosis %>% filter (is.na(deathdate))
head(osteoporosis)
```

```{r}
hypoxemia<-dbGetQuery(db5_con, "SELECT * FROM hypo")
#hypoxemia <- hypoxemia %>% filter (is.na(deathdate))
head(hypoxemia)

```

```{r}
hypertension<-dbGetQuery(db6_con, "SELECT * FROM hypertension")
hypertension <- hypertension %>% filter (is.na(deathdate))
head(hypertension)
```

```{r}
acute_bronchitis<-dbGetQuery(db7_con, "SELECT * FROM acute_bronchitis")
# acute_bronchitis <- acute_bronchitis %>% filter (is.na(deathdate))
head(acute_bronchitis)
```

```{r}
anemia<-dbGetQuery(db8_con, "SELECT * FROM anemia")
# acute_bronchitis <- acute_bronchitis %>% filter (is.na(deathdate))
head(anemia)
```

```{r}
heart_failure<-dbGetQuery(db9_con, "SELECT * FROM heart_failure")
heart_failure <- heart_failure %>% filter (is.na(deathdate))
head(heart_failure)
```

### 3) Data Analysis


#### i) Number of Patients 

```{r}
# Count of distinct IDs from each databases. 

pneumonia_count <- pneumonia %>% dplyr::distinct(id) %>% summarise(number_of_patients=n())
pneumonia_count$condition <- 'pneumonia'
viral_sinusitis_count <- viral_sinusitis %>% dplyr::distinct(id) %>% summarise(number_of_patients=n())
viral_sinusitis_count$condition <- 'viral sinusitis'
chronic_sinusitis_count <- chronic_sinusitis %>% dplyr::distinct(patient) %>% summarise(number_of_patients=n())
chronic_sinusitis_count$condition <- 'chronic sinusitis'
osteoporosis_count <- osteoporosis %>% dplyr::distinct(patient) %>% summarise(number_of_patients=n())
osteoporosis_count$condition <- 'osteoporosis'
hypoxemia_count <- hypoxemia %>% dplyr::distinct(id) %>% summarise(number_of_patients=n())
hypoxemia_count$condition <- 'hypoxemia'
hypertension_count <- hypertension %>% dplyr::distinct(id) %>% summarise(number_of_patients=n())
hypertension_count$condition <- 'hypertension'
acute_bronchitis_count <- acute_bronchitis %>% dplyr::distinct(id) %>% summarise(number_of_patients=n())
acute_bronchitis_count$condition <- 'acute_bronchitis'
anemia_count <- anemia %>% dplyr::distinct(patient) %>% summarise(number_of_patients=n())
anemia_count$condition <- 'anemia'
heart_failure_count <- heart_failure %>% dplyr::distinct(id) %>% summarise(number_of_patients=n())
heart_failure_count$condition <- 'heart_failure'

combined_count <- rbind(pneumonia_count,viral_sinusitis_count,chronic_sinusitis_count, osteoporosis_count,hypoxemia_count,hypertension_count,acute_bronchitis_count,anemia_count, heart_failure_count)

# Ordered from high to low. 
combined_count <- combined_count %>% arrange (desc(number_of_patients))
print(combined_count)
```

Visualization of patient count. 

```{r}
ggplot(combined_count, aes(x = condition, y = number_of_patients, fill = condition)) +
  geom_bar(stat = "identity") +
  labs(x=NULL, y = "Number of Patients", title = "Bar Graph of Number of Patients by Condition") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#### ii) Age of Patients

```{r}
# Age in pneumonia data
pneumonia <- pneumonia %>%
  mutate(age=ceiling(as.numeric(difftime(Sys.Date(), birthdate))/365))

#print(head(pneumonia))
pneumonia_age <- pneumonia$age
print(head(pneumonia_age))
```

```{r}
# Age in viral sinusitis
viral_sinusitis_age <- viral_sinusitis$age
print(head(viral_sinusitis_age))
```

```{r}
# Age in chronic sinusitis data
chronic_sinusitis <- chronic_sinusitis %>%
  mutate(age=ceiling(as.numeric(difftime(Sys.Date(), birthdate))/365))

#print(head(chronic_sinusitis))
chronic_sinusitis_age <- chronic_sinusitis$age
print(head(chronic_sinusitis_age))
```

```{r}
# Age in osteoporosis data
osteoporosis <- osteoporosis %>%
  mutate(age=ceiling(as.numeric(difftime(Sys.Date(), birthdate))/365))

#print(head(osteoporosis))
osteoporosis_age <- osteoporosis$age
#print(head(osteoporosis_age)
print(range(osteoporosis_age))
```

```{r}
# Age in hypertension data
hypertension <- hypertension %>%
  mutate(age=ceiling(as.numeric(difftime(Sys.Date(), birthdate))/365))

#print(head(hypertension))
hypertension_age <- hypertension$age
print(head(hypertension_age))
```

```{r}
# Age in heart failure
heart_failure <- heart_failure %>%
  mutate(age=ceiling(as.numeric(difftime(Sys.Date(), birthdate))/365))

#print(head(heart_failure))
heart_failure_age <- heart_failure$age
print(head(heart_failure_age))
```

Visualization of patient age distribution for each condition

```{r}
# Calculate histograms for each dataset
hist1 <- hist(pneumonia_age, plot = FALSE)
hist2 <- hist(viral_sinusitis_age, plot = FALSE)
hist3 <- hist(chronic_sinusitis_age, plot = FALSE)
hist4 <- hist(osteoporosis_age, plot = FALSE)
hist5 <- hist(hypertension_age, plot = FALSE)
hist6 <- hist(heart_failure_age, plot = FALSE)

# Get the maximum count across all datasets
max_count <- max(c(hist1$counts, hist2$counts, hist3$counts,hist4$counts,hist5$counts))

# Create plot with adjusted margins
par(mar = c(3, 3, 2, 1))

# Plot counts as lines
plot(hist1$breaks[-1], hist1$counts, type = "l", col = "blue", main = "Combined Line Plot of Counts", xlab = "Value", ylab = "Count", xlim = range(c(hist1$breaks, hist2$breaks, hist3$breaks,hist4$breaks,hist5$breaks)), ylim = c(0, max_count))
lines(hist2$breaks[-1], hist2$counts, col = "red")
lines(hist3$breaks[-1], hist3$counts, col = "green")
lines(hist4$breaks[-1], hist4$counts, col = "orange")
lines(hist5$breaks[-1], hist5$counts, col = "purple")
lines(hist6$breaks[-1], hist6$counts, col = "skyblue")

# Add legend
legend("topright", legend = c("penumonia", "viral sinusitis","chronic sinusitis", "osteoporosis","hypertension","heart failure"), col = c("blue", "red", "green","orange","purple","skyblue"), lty = 1)


```


Average, Minimum and Maximum Age in Patient Groups

```{r}
pneumonia_avgage <- pneumonia %>% summarise(
    avg_age = mean(age, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE)
  )
pneumonia_avgage$name <- 'pneumonia'
viral_sinusitis_avgage <- viral_sinusitis %>% summarise(
    avg_age = mean(age, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE)
  )
viral_sinusitis_avgage$name <- 'viral sinusitis'
chronic_sinusitis_avgage <- chronic_sinusitis %>% summarise(
    avg_age = mean(age, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE)
  )
chronic_sinusitis_avgage$name <- 'chronic sinusitis'
osteoporosis_avgage <- osteoporosis %>% summarise(
    avg_age = mean(age, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE)
  )
osteoporosis_avgage$name <- 'osteoporosis'
hypertension_avgage <- hypertension %>% summarise(
    avg_age = mean(age, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE)
  )
hypertension_avgage$name <- 'hypertension'
heart_failure_avgage <- heart_failure %>% summarise(
    avg_age = mean(age, na.rm = TRUE),
    min_age = min(age, na.rm = TRUE),
    max_age = max(age, na.rm = TRUE)
  )
heart_failure_avgage$name <- 'heart_failure'


combined_age <- rbind(pneumonia_avgage,viral_sinusitis_avgage,chronic_sinusitis_avgage,osteoporosis_avgage,hypertension_avgage,heart_failure_avgage)



print(combined_age)


```


#### iii) Gender of Patients

```{r}

# Function to calculate male and female percentages
calculate_percentages <- function(df) {
  # Count occurrences of male and female
  n_male <- sum(df$gender == "M")
  n_female <- sum(df$gender == "F")
  
  # Calculate total count
  total <- n_male + n_female
  
  # Check for division by zero (no data)
  if (total == 0) {
    return(data.frame(male = NA, female = NA))
  } else {
    # Calculate percentages
    male_pct <- round(n_male / total * 100, 2)
    female_pct <- round(n_female / total * 100, 2)
    
    # Return data frame with percentages
    return(data.frame(male = male_pct, female = female_pct))
  }
}

# Calculate percentages for each dataframe
pneumonia_gender <- calculate_percentages(pneumonia)
viral_sinusitis_gender <- calculate_percentages(viral_sinusitis)
chronic_sinusitis_gender <- calculate_percentages(chronic_sinusitis)
hypoxemia_gender <- calculate_percentages(hypoxemia)
osteoporosis_gender <- calculate_percentages(osteoporosis)
hypertension_gender <- calculate_percentages(hypertension)
heart_failure_gender <- calculate_percentages(heart_failure)

# Combine percentages into a new dataframe
combined_percentages <- rbind(pneumonia_gender, viral_sinusitis_gender,chronic_sinusitis_gender,hypoxemia_gender, osteoporosis_gender,hypertension_gender,heart_failure_gender)

combined_percentages$condition <- c("Pneumonia", "Viral Sinusitis", "Chronic Sinusitis","Hypoxemia", "Osteoporosis","Hypertension","Heart Failure")

print(combined_percentages)

```


#### iv) Race of Patients

```{r}
#First check unique values
unique_values <- unique (pneumonia$race)
unique_values

```

```{r}

# Function to calculate race percentages
calculate_racepercentages <- function(df) {
  # Count occurrences of races
  n_white <- sum(df$race == "white")
  n_asian <- sum(df$race == "asian")
  n_black <- sum(df$race == "black")
  n_native <- sum(df$race == 'native')
  n_other <- sum(df$race == 'other')
  
  # Calculate total count
  total <- n_white + n_asian + n_black + + n_native + n_other
  white_pct <- round(n_white / total * 100, 2)
  asian_pct <- round(n_asian / total * 100, 2)
  black_pct <- round(n_black / total * 100, 2)
  native_pct <- round(n_native / total * 100, 2)
  other_pct <- round(n_other / total * 100, 2)
    
    # Return data frame with percentages
    return(data.frame(white = white_pct, black = black_pct, asian=asian_pct, native=native_pct, other=other_pct))
  }

# Calculate percentages for each dataframe
pneumonia_race <- calculate_racepercentages(pneumonia)
viral_sinusitis_race <- calculate_racepercentages(viral_sinusitis)
chronic_sinusitis_race <- calculate_racepercentages(chronic_sinusitis)
hypoxemia_race <- calculate_racepercentages(hypoxemia)
osteoporosis_race <- calculate_racepercentages(osteoporosis)
hypertension_race <- calculate_racepercentages(hypertension)
heart_failure_race <- calculate_racepercentages(heart_failure)

# Combine percentages into a new dataframe
combined_racepercentages <- rbind(pneumonia_race, viral_sinusitis_race,chronic_sinusitis_race, hypoxemia_race, osteoporosis_race,hypertension_race)
combined_racepercentages$condition <- c("Pneumonia", "Viral Sinusitis","Chronic Sinusitis","Hypoxemia", "Osteoporosis","Hypertension")


print(combined_racepercentages)

```